% Conclusion
\section{Conclusion}

The age of the slow, deliberative robot is over. The speed of modern warfare—defined by hypersonics, swarms, and directed energy—has compressed the OODA loop beyond the physiological limits of human cognition. In this new era, the insistence on ``Human-in-the-loop'' control is not a moral safeguard; it is a suicide pact. We must automate.

However, automation without ethics is merely efficient slaughter. This paper has argued that existing frameworks—from Asimov's literary plot devices to the DoD's vague requirements for judgment—are insufficient for the specific domain of defensive AI. We proposed the ``Sentinel'' Ruleset: a hierarchical, lexicographically ordered ethical engine that priorities \textit{Distinction} over \textit{Obedience} and \textit{Human Life} over \textit{Material Assets}.

\subsection{Summary of Contributions}
Our contribution is threefold:
\begin{enumerate}
    \item \textbf{Defensive Specificity:} We isolated ``Defensive AI'' as a distinct moral category, separate from offensive LAWS, governed by \textit{Jus ad Vim} and strict Neutralization.
    \item \textbf{Operationalized Ethics:} We translated abstract IHL principles into verifiable engineering constraints (e.g., Fail Open architectures, 99.9\% Confidence Intervals).
    \item \textbf{The Martyrdom Paradigm:} We established the moral obligation of the autonomous system to sacrifice itself—and the assets it protects—to strictly minimize human collateral damage, rejecting the doctrine of Force Protection for machines.
\end{enumerate}

\subsection{Future Work}
The path from this theoretical framework to a deployable system requires urgent interdisciplinary work.
\textbf{Simulation:} The next phase of research must involve high-fidelity simulations (e.g., in Unity or Ansys) to stress-test these rules against adversarial inputs.
\textbf{Legal Engineering:} We call for a formalization of ``Machine Readable IHL''—a digital Geneva Convention that can be parsed by autonomous agents.
\textbf{Policy Standards:} We urge NATO and allied defense bodies to adopt the ``Sentinel Standard'' as a baseline for the certification of defensive autonomous systems.

In the end, the Sentinel is not a soldier. It has no honor, no courage, and no instinct for survival. It is a shield. And like any shield, its only purpose is to break so that the human behind it does not have to.
